\documentclass[10pt,fleqn]{article}
\usepackage{standard}

\geometry{left=27mm,right=27mm,top=40mm,bottom=30mm}
\allowdisplaybreaks

\newcommand{\exerciseHeader}{%
  \hrule
  \begin{center}
    \Large
    \scshape
    Advanced Quantum Theory \\ Exercise Sheet 11
  \end{center}
  \medskip
  {
    \footnotesize
    \begin{minipage}[c]{0.49\textwidth}
      Markus Pawellek \\
      markuspawellek@gmail.com
    \end{minipage}
    \hfill
    \begin{minipage}[c]{0.49\textwidth}
      \raggedleft
      \today
    \end{minipage}
  }
  \medskip
  \hrule
  \bigskip
}

\newcommand{\bra}[1]{\left\langle #1 \right\vert}
\newcommand{\ket}[1]{\left\vert #1 \right\rangle}
\newcommand{\bracket}[2]{\left\langle #1 \middle\vert #2 \right\rangle}
\DeclareMathOperator{\derivative}{\mathrm{D}}

\begin{document}
  \selectlanguage{english}

  \exerciseHeader

  \section*{Problem 18: Perturbation Theory} % (fold)
  \label{sec:problem_18_perturbation_theory}
    \begin{multicols}{2}
      Let $\roundBrackets{\mathscr{H},\bracket{\cdot}{\cdot}}$ be a separable Hilbert space which is modeling the perturbed and the unperturbed system.
      For the formal definition of both systems we will define $\mathscr{I}$ as a countable index set with the same cardinality as an orthonormal basis of $\mathscr{H}$.

      The Hamiltonian $\function{H}{\mathscr{H}}{\mathscr{H}}$ of the unperturbed system is assumed to have a discrete spectrum of real Energies.
      \[
        σ(H) = \set{E_n}{n\in\mathscr{I}}\subset\setReal
      \]
      For all $n\in\mathscr{I}$ we denote $\ket{n}\in\mathscr{H}$ to be the unique normalized eigenstate of $H$ with the eigenvalue $E_n$.
      \[
        H\ket{n} = E_n\ket{n}
        \separate
        \bracket{n}{n} = 1
      \]
      \[
        \forall m\in\mathscr{I},m\neq n: \quad E_m\neq E_n
      \]
      Based on this and due to the self-adjointness of the Hamiltonian $H$ we can derive that
      % , for all $n,m\in\mathscr{I}$ with $n\neq m$ we can derive that $\ket{m}$ and $\ket{n}$ are orthogonal.
      % \[
      %   \bra{m}H\ket{n} = E_n \bracket{m}{n} = E_m \bracket{m}{n}
      % \]
      % \[
      %   \implies (E_n-E_m) \bracket{m}{n} = 0
      % \]
      % \[
      %   \implies \bracket{m}{n} = 0
      % \]
      % Hence, for all $n,m\in\mathscr{I}$ we can state the following.
      % \[
      %   \bracket{m}{n} = δ_{mn}
      % \]
      $H$ itself is non-degenerate and that the set $\set{\ket{n}}{n\in\mathscr{I}}$ is building an orthonormal basis of the Hilbert space $\mathscr{H}$.

      Further, we describe the perturbation of the system by the self-adjoint operator $\function{V}{\mathscr{H}}{\mathscr{H}}$.
      For the application of $V$ to $H$ we will use an interaction strength parameter which is introduced by using a function $\tilde{H}$ as shown below.
      \[
        \function{\tilde{H}}{[0,1]}{\mathrm{L}(\mathscr{H},\mathscr{H})}
        \separate
        \tilde{H}(λ) \define H + λ V
      \]
      The continuous parameterization will enable us to analyze the solution of the perturbed system $\tilde{H}(λ)$ in form of a series expansion in terms of λ.
      Please note that $\tilde{H}(1)$ is describing the system under the full perturbation.

      We will assume that the contribution of $V$ is small and therefore not changes the essential properties.
      Hence, we choose $λ\in[0,1]$ to be arbitrary.
      $\tilde{H}(λ)$ has a discrete spectrum.
       % of non-degeneracy and that of a discrete spectrum of $\tilde{H}(λ)$ for all $λ\in[0,1]$.
      % Let $n\in\mathscr{I}$ be arbitrary.
      \[
        σ\roundBrackets{\tilde{H}(λ)} = \set{E_n(λ)}{n\in\mathscr{I}} \subset \setReal
      \]
      $\tilde{H}(λ)$ is non-degenerate and we again assume its eigenstates to be normalized and therefore to build an orthonormal eigenbasis with respect to $\tilde{H}(λ)$
      \footnote{%
        The non-degeneracy of $\tilde{H}(λ)$ for all $λ\in[0,1]$ is not used explicitly.
        Therefore the derived formulas should be valid even if the perturbation described by $V$ introduces degeneracies into the non-degenerate system described by $H$.%
      }.
      So for all $n\in\mathscr{I}$ it holds that
      \[
        \tilde{H}(λ) \ket{n(λ)} = E_n(λ) \ket{n(λ)}
      \]
      \[
        \forall m\in\mathscr{I},m\neq n: \quad E_m(λ)\neq E_n(λ)
      \]
      \[
        \bracket{n(λ)}{n(λ)} = 1
      \]
      Additionally, we assume that we are able to expand $E_n(λ)$ and $\ket{n(λ)}$ in terms of $λ$ for all $λ\in[0,1]$ for some constant coefficients.
      \[
        \set{E_{nk}}{}_{k=0}^\infty \subset \setReal
        \separate
        E_n(λ) = \sum_{k=0}^\infty λ^k E_{nk}
      \]
      \[
        \set{\ket{n_k}}{}_{k=0}^\infty\subset\mathscr{H}
        \separate
        \ket{n(λ)} = \sum_{k=0}^\infty λ^k \ket{n_k}
      \]

      Taking the series expansions one can now derive inductive formulas for the coefficients.
      For that let $n\in\mathscr{I}$ and $λ\in[0,1]$ be arbitrary.
      We will start with the left-hand side of the eigenvalue equation.
      \begin{align*}
        &\tilde{H}(λ)\ket{n(λ)}\\
        &= (H+λV)\sum_{k=0}^\infty λ^k \ket{n_k} \\
        &= \sum_{k=0}^\infty λ^k H\ket{n_k} + \sum_{k=0}^\infty λ^{k+1} V\ket{n_k} \\
        &= \sum_{k=0}^\infty λ^{k} H\ket{n_k} + \sum_{k=1}^\infty λ^{k} V\ket{n_{k-1}} \\
        &= H\ket{n_0} + \sum_{k=1}^\infty λ^{k} \roundBrackets{H\ket{n_k} + V\ket{n_{k-1}}}
      \end{align*}
      For the right-hand side of the eigenvalue equation we do get the following.
      \begin{align*}
        &E_n(λ)\ket{n(λ)} \\
        &= \roundBrackets{\sum_{k=0}^\infty λ^k E_{nk}}\roundBrackets{\sum_{k=0}^\infty λ^k \ket{n_k}} \\
        &= \sum_{p,q=0}^\infty λ^{p+q} E_{np}\ket{n_q} \\
        &= \sum_{k=0}^\infty λ^k \sum_{p=0}^k E_{np}\ket{n_{k-p}}
      \end{align*}
      Looking at the eigenvalue equation as a whole we are now able to do a comparison of coefficients.
      This results in two equations.
      One for the starting values and one for all $k\in\setNatural$.
      \[
        H\ket{n_0} = E_{n0}\ket{n_0}
      \]
      \[
        H\ket{n_k} + V\ket{n_{k-1}} = \sum_{p=0}^k E_{np}\ket{n_{k-p}}
      \]
      After inserting the series expansions we can do something similar for the normalization condition.
      \begin{align*}
        1 &= \bracket{n(λ)}{n(λ)} \\
        &= \sum_{p,q=0} λ^{p+q} \bracket{n_p}{n_q} \\
        &= \sum_{k=0}^\infty λ^k \sum_{p=0}^k \bracket{n_p}{n_{k-p}}
      \end{align*}
      Again we do comparison of coefficients and get two equations.
      One for the starting value and one for all $k\in\setNatural$.
      \[
        \bracket{n_0}{n_0} = 1
      \]
      \begin{align*}
        \bracket{n_0}{n_k} + \bracket{n_k}{n_0}
        &= 2 \Re(\bracket{n_0}{n_k}) \\
        &= -\sum_{p=1}^{k-1} \bracket{n_p}{n_{k-p}}
      \end{align*}
      To make the second equation simpler consider that the overall phase is not determined in quantum mechanics.
      Hence, without loss of generality, we may assume $\bracket{n_0}{n_k}=\Re(\bracket{n_0}{n_k})\in\setReal$ is purely real.
      For all $k\in\setNatural$ the second equation becomes the following.
      \[
        \bracket{n}{n_k} = -\frac{1}{2}\sum_{p=1}^{k-1} \bracket{n_p}{n_{k-p}}
      \]

      After deriving these formulas we will first take a look at the starting value equations.
      \[
        H\ket{n_0} = E_{n0}\ket{n_0}
        \separate
        \bracket{n_0}{n_0} = 1
      \]
      We see that $\ket{n_0}$ is a normalized eigenfunction of $H$ with the eigenvalue $E_{n0}$.
      Therefore we can state the following.
      \[
        E_{n0} \in σ(H)
        \separate
        \ket{n_0} \in \set{\ket{m}}{m\in\mathscr{I}}
      \]
      We are free to choose the specific eigenfunction because this is the freedom of permuting the eigenstates.
      But for consistency we will use the straightforward definition.
      \[
        \ket{n_0} \define \ket{n}
        \separate
        E_{n0} \define E_n
      \]

      Now we take the inductive formula of the eigenvalue equation and insert the starting values to get an operator equation for $\ket{n_k}$ for all $k\in\setNatural$.
      \[
        \roundBrackets{H-E_n}\ket{n_k} = -V\ket{n_{k-1}} + \sum_{p=1}^{k} E_{np}\ket{n_{k-p}}
      \]
      But this equation gives us some information about the energy shift $E_{nk}$ as well.
      For that we will apply $\bra{n}$ on the equation for all $k\in\setNatural$.
      \begin{align*}
        0
        &= \bra{n}\roundBrackets{H-E_n}\ket{n_k} \\
        &= -\bra{n}V\ket{n_{k-1}} + \sum_{p=1}^k E_{np}\bracket{n}{n_{k-p}}
      \end{align*}
      By solving the equation for $E_{nk}$ one gets the following for all $k\in\setNatural$.
      \[
        E_{nk} = \bra{n}V\ket{n_{k-1}} - \sum_{p=1}^{k-1} E_{np}\bracket{n}{n_{k-p}}
      \]
      For the eigenstates $\ket{n_k}$ we basically have to invert the operator $H-E_n$.
      But due to its singularity this is not possible.
      For that reason we will first express $\ket{n_k}$ in terms of the orthonormal eigenbasis of $\mathscr{H}$ with respect to $H$.
      \begin{align*}
        \ket{n_k}
        &= \sum_{m\in\mathscr{I}} \bracket{m}{n_k} \ket{m} \\
        &= \bracket{n}{n_k}\ket{n} + \sum_{\substack{m\in\mathscr{I}\\m\neq n}} \bracket{m}{n_k} \ket{m}
      \end{align*}
      Let now $m\in\mathscr{I}$ with $m\neq n$ be arbitrary as well and apply $\bra{m}$ on the operator equation for $\ket{n_k}$ for all $k\in\setNatural$.
      \begin{align*}
        &\bra{m}\roundBrackets{H-E_n}\ket{n_k} \\
        &= \roundBrackets{E_m-E_n}\bracket{m}{n_k} \\
        &= -\bra{m}V\ket{n_{k-1}} + \sum_{p=1}^k E_{np}\bracket{m}{n_{k-p}} \\
        &= -\bra{m}V\ket{n_{k-1}} + \sum_{p=1}^{k-1} E_{np}\bracket{m}{n_{k-p}}
      \end{align*}
      Solving this for $\bracket{m}{n_k}$ for all $k\in\setNatural$ gives us the following equations.
      \[
        \bracket{m}{n_k} = -\frac{\bra{m}V\ket{n_{k-1}}}{E_m-E_n} + \sum_{p=1}^{k-1} \frac{E_{np}\bracket{m}{n_{k-p}}}{E_m - E_n}
      \]
      Inserting now the equations for $\bracket{n}{n_k}$ and $\bracket{m}{n_k}$ into the expansion with respect to the eigenbasis yields the following for all $k\in\setNatural$.
      \begin{align*}
        \ket{n_k} &= -\frac{1}{2}\sum_{p=1}^{k-1} \bracket{n_p}{n_{k-p}}\ket{n} \\
        &+ \sum_{\substack{m\in\mathscr{I}\\m\neq n}}
          \begin{aligned}[t]
            &\left[ -\frac{\bra{m}V\ket{n_{k-1}}}{E_m-E_n} \right. \\
            &\left. + \sum_{p=1}^{k-1} \frac{E_{np}\bracket{m}{n_{k-p}}}{E_m - E_n} \right] \ket{m}
          \end{aligned}
      \end{align*}

      With this last equation we now have obtained explicit inductive formulas with their respective starting values for $E_{nk}$ and $\ket{n_k}$ for all $k\in\setNatural$.
      Through an iterative procedure beginning by $k=1$ one can now directly obtain the formulas for the energy and state shifts.
      \[
        E_{n1} = \bra{n}V\ket{n}
      \]
      \[
        \ket{n_1} = - \sum_{\substack{m\in\mathscr{I} \\ m\neq n}} \frac{\bra{m}V\ket{n}}{E_m-E_n} \ket{m}
      \]
      The same formulas will be used for $k=2$.
      But this time we will directly insert the equations for $E_{n1}$ and $\ket{n_1}$ obtained above.
      \begin{align*}
        E_{n2}
        &= \bra{n}V\ket{n_1} - E_{n1}\bracket{n}{n_1} \\
        &= -\sum_{\substack{m\in\mathscr{I} \\ m\neq n}} \frac{\bra{m}V\ket{n}}{E_m-E_n} \bra{n}V\ket{m}
      \end{align*}
      \begin{align*}
        \ket{n_2}
        &= -\frac{1}{2} \bracket{n_1}{n_{1}}\ket{n} \\
        &+ \sum_{\substack{m\in\mathscr{I}\\m\neq n}}
          \begin{aligned}[t]
            &\left[ -\frac{\bra{m}V\ket{n_{1}}}{E_m-E_n} \right. \\
            &\left. + \frac{E_{n1}\bracket{m}{n_{1}}}{E_m - E_n} \right] \ket{m}
          \end{aligned}
        \\
        &= -\frac{1}{2} \sum_{\substack{m,k\in\mathscr{I}\\m,k\neq n}} \frac{\bra{n}V\ket{m}}{E_m-E_n} \frac{\bra{k}V\ket{n}}{E_k-E_n} \bracket{m}{k} \ket{n} \\
        &+ \sum_{\substack{m\in\mathscr{I}\\m\neq n}}
          \begin{aligned}[t]
            &\left[ \sum_{\substack{k\in\mathscr{I} \\ k\neq n}} \frac{\bra{k}V\ket{n}}{E_k-E_n} \frac{\bra{m}V\ket{k}}{E_m-E_n} \right. \\
            &\left. - \sum_{\substack{k\in\mathscr{I} \\ k\neq n}} \frac{\bra{k}V\ket{n}}{E_k-E_n} \frac{\bra{n}V\ket{n} \bracket{m}{k}}{E_m - E_n} \right] \ket{m}
          \end{aligned}
        \\
        &= -\frac{1}{2} \sum_{\substack{m\in\mathscr{I}\\m\neq n}} \frac{\absolute{\bra{m}V\ket{n}}^2}{(E_m-E_n)^2} \ket{n} \\
        &+ \sum_{\substack{m,k\in\mathscr{I}\\m,k\neq n}} \frac{\bra{k}V\ket{n} \bra{m}V\ket{k}}{(E_k-E_n)(E_m-E_n)} \ket{m} \\
        &- \sum_{\substack{m\in\mathscr{I}\\m\neq n}} \frac{\bra{n}V\ket{n} \bra{m}V\ket{n}}{(E_m-E_n)^2} \ket{m}
      \end{align*}
      This proves the proposition.
      \hfill$\square{}$
    \end{multicols}
  % section problem_18_perturbation_theory (end)

  \newpage

  \section*{Problem 19: Scattering off a Radial Potential} % (fold)
  \label{sec:problem_19}
    \begin{multicols}{2}
      \paragraph{Preliminaries:}
      Let $\roundBrackets{\mathscr{H},\bracket{\cdot}{\cdot}}$ be the separable Hilbert space of three-dimensional square-integrable functions with Lebesgue-measure λ and the typical scalar product which models the scattering of a nonrelativistic particle of mass $m\in\setReal^+$.
      \[
        \mathscr{H}
        % \define \mathscr{L}^2\roundBrackets{\setReal^3}
        = \set{\function{φ}{\setReal^3}{\setComplex}}{\integral{\setReal^3}{}{\absolute{φ}^2}{λ} < \infty}
      \]
      \[
        \function{\bracket{\cdot}{\cdot}}{\mathscr{H}}{\mathscr{H}}
        \separate
        \bracket{φ}{ψ} \define \integral{\setReal^3}{}{\bar{φ}ψ}{λ}
      \]
      Let $V\in\mathscr{H}$ be the spherically symmetric potential of the given system.
      We will use the spherical standard parameterization $(M,μ)$ of $\setReal^3$.
      \[
        M\define \setReal^+\times[0,π]\times[0,2π)
      \]
      \[
        \function{μ}{M}{\setReal^3}
        \separate
        μ(r,ϑ,φ) \define
        \begin{pmatrix}
          r\sinϑ\cosφ \\
          r\sinϑ\sinφ \\
          r\cosϑ
        \end{pmatrix}
      \]
      The potential $V$ can be rewritten in form of a scaled radial potential $U$.
      \[
        \function{U}{\setReal^+}{\setReal}
      \]
      \[
        \forall x\in\setReal^3:\quad U(\norm{x}) \define \frac{2m}{\hbar^2}V(x)
      \]
      Hence, the typical Hamilton operator is given by the expression below.
      \[
        \function{H}{\mathscr{H}}{\mathscr{H}}
        \separate
        H \define -\frac{\hbar^2}{2m}\laplacian + V
      \]

      \paragraph{Expansion:}
      Let $k\in\setReal^+$ be arbitrary and let $\ket{ψ_k}\in\mathscr{H}$ be a solution of the Schrödinger equation with eigenvalue $E(k)$.
      \[
        H\ket{ψ_k} = E(k)\ket{ψ_k}
        \separate
        E(k)\define \frac{\hbar^2k^2}{2m}
      \]
      By inserting the definitions and multiplying with $2m\hbar^{-2}$ this equation takes on the following form.
      \[
        \roundBrackets{\laplacian - U\circ\norm{\cdot}  + k^2}\ket{ψ_k} = 0
      \]
      To model the scattering of a particle, we assume the particle to be a plane wave in $z$-direction.
      Because of that and due to the radial potential the wave function should not depend on the azimuth angle.
      So for all $(r,ϑ,φ)\in M$ we can reformulate the wave function as follows.
      \[
        \tilde{ψ}_k(r,ϑ) \define ψ_k\circ μ(r,ϑ,φ)
      \]
      Thanks to the spherically symmetric potential, we can expand $\ket{ψ_k}$ in terms of radial functions $\function{A_l}{\setReal^+}{\setComplex}$ and spherical harmonics $\mathscr{Y}_{l0}$ with coefficients $c_{l}\in\setComplex$ for all $l\in\setNatural_0$ and $r\in\setReal^+$.
      Therefore, we get the following for all $(r,ϑ,φ)\in M$.
      \begin{align*}
        \tilde{ψ}_k(r,ϑ)
        &= \sum_{l=0}^\infty c_{l} A_l(r) \mathscr{Y}_{l0}(ϑ) \\
        &= \sum_{l=0}^\infty \tilde{c}_l A_l(r) P_l(\cos ϑ)
      \end{align*}
      Here $P_l$ are the Legendre polynomials and $\tilde{c}_l\in\setComplex$ for all $l\in\setNatural_0$.
      For the expansion to be valid, the radial functions $A_l$ have to fulfill the radial Schrödinger equation for all $l\in\setNatural_0$ and $r\in\setReal^+$.
      \[
        \begin{aligned}
          0 =&\ r^2 A''_l(r) + 2r A'_l(r) \\
          &+ \boxBrackets{\roundBrackets{k^2 - U(r)}r^2 - l(l+1)} A_l(r)
        \end{aligned}
      \]
      This equation can be simplified by defining a variant of the radial function for all $l\in\setNatural_0$ and $r\in\setReal^+$.
      \[
        \function{u_l}{\setReal^+}{\setComplex}
        \separate
        u_l(r) := r A_l(r)
      \]
      Then for all $l\in\setNatural_0$ the variants $u_l$ fulfill the following differential equation for all $r\in\setReal^+$.
      \[
        u_l''(r) + \boxBrackets{k^2-U(r)-\frac{l(l+1)}{r^2}}u_l(r) = 0
      \]

      \paragraph{Differential Equation:}
      In the expansion of the final-state wave function of the scattered particle $ψ_k$ takes on the following form for all $r\in\setReal^+$ and $ϑ\in[0,π]$.
      \[
        ψ_k(r,ϑ) = \sum_{l=0}^\infty i^l (2l+1) P_l(\cos ϑ) \frac{e^{iδ_l}}{kr}g_l(r)
      \]
      Here $δ_l\in[0,2π]$ is the phase shift and $\function{g_l}{\setReal^+}{\setComplex}$ another radial function for all $l\in\setNatural_0$.
      Comparing this expression with the former wave expansion we find a proportional relation for all $l\in\setNatural_0$ and $r\in\setReal^+$.
      \[
        \frac{g_l(r)}{r} \propto A_l(r)
        \quad\implies\quad
        g_l(r) \propto u_l(r)
      \]
      For any $l\in\setNatural_0$ both differential equations for $A_l$ and for $u_l$ are linear homogeneous differential equations of second order.
      Hence, any solution of these equations multiplied by any constant is also a solution.
      As a result $g_l$ has to fulfill the same differential equation as the variants $u_l$ for all $l\in\setNatural_0$ and $r\in\setReal^+$.
      \[
        g_l''(r) + \boxBrackets{k^2-U(r)-\frac{l(l+1)}{r^2}}g_l(r) = 0
      \]

      \paragraph{Asymptotic Behavior:}
      Let $l\in\setNatural_0$ be arbitrary.
      To get a general idea of what is happening near infinity we will first look at the asymptotic behavior of the differential equation fulfilled by $g_l$.
      We will assume that $U(r)\converges[r\to\infty] 0$ which leads us to the following equation.
      Choose $r\in\setReal^+$ to be near infinity.
      \[
        g_l''(r) + k^2 g_l(r) = 0
      \]
      This is the differential equation of the classical harmonic oscillator and can be solved easily for constants $A,B\in\setComplex$.
      \[
        g_l(r) = A e^{ikr} + B e^{-ikr}
      \]
      Therefore the variants of the radial functions will asymptotically behave the same way as a spherical wave.
      To determine the values $A$ and $B$ we will use another more detailed approach based on the scattering amplitude.

      The scattering amplitude $\function{f_k}{[0,π]}{\setComplex}$ in its partial wave expansion with partial-wave amplitudes $f_{kl}\in\setComplex$ for $l\in\setNatural_0$ can be written as follows for all $ϑ\in[0,π]$.
      \[
        f_k(ϑ) = \sum_{l=0}^\infty (2l+1)f_{kl} P_l(\cos ϑ)
      \]
      Applying this formula on the wave function of the scattered particle we get the following.
      \[
        ψ_k(r) \converges
        \begin{aligned}[t]
          & \sum_{l=0}^\infty (2l+1) P_l(\cos ϑ) \\
          & \cdot \frac{1}{2ik} \boxBrackets{\roundBrackets{1+2ikf_{kl}} \frac{e^{ikr}}{r} - \frac{e^{-i(kr-lπ)}}{r}}
        \end{aligned}
      \]
      By definition, the partial-wave amplitudes are related to the phase shifts.
      \[
        1+2ikf_{kl} = e^{2iδ_l}
      \]
      Taking this equation and comparing the partial-wave expansion based on the scattering amplitude to the already given expansion we can now conclude the following.
      \[
        \begin{aligned}
          i^l &e^{iδ_l} \frac{g_l(r)}{kr}
          \longrightarrow \frac{1}{2ik} \boxBrackets{e^{2iδ_l} \frac{e^{ikr}}{r} - \frac{e^{-i(kr-lπ)}}{r}} \\
          &= \frac{e^{i\roundBrackets{δ_l+\frac{π}{2}l}}}{2ikr} \boxBrackets{e^{i\roundBrackets{kr - \frac{π}{2}l + δ_l}} - e^{-i\roundBrackets{kr - \frac{π}{2}l + δ_l}}} \\
          &=  e^{il\frac{π}{2}} \frac{e^{iδ_l}}{kr} \sin\roundBrackets{ kr -\frac{π}{2}l + δ_l } \\
          &= i^l \frac{e^{iδ_l}}{kr} \sin\roundBrackets{ kr -\frac{π}{2}l + δ_l }
        \end{aligned}
      \]
      The detailed asymptotic behavior of $g_l$ is explicitly be described below.
      \[
        g_l(r) \converges \sin\roundBrackets{ kr -\frac{π}{2}l + δ_l }
      \]

      \paragraph{Wronskian Determinant:}
      Let $V_1$ and $V_2$ be potentials with the same properties as $V$ and define $U_1$, $δ_l^{(1)}$ and $U_2$, $δ_l^{(2)}$ analog to $U$ and $δ_l$.
      Let $\function{g_l,h_l}{\setReal^+}{\setComplex}$ be radial functions such that they fulfill the following differential equations for all $r\in\setReal^+$.
      \[
        g_l''(r) + \boxBrackets{k^2-U_1(r)-\frac{l(l+1)}{r^2}}g_l(r) = 0
      \]
      \[
        h_l''(r) + \boxBrackets{k^2-U_2(r)-\frac{l(l+1)}{r^2}}h_l(r) = 0
      \]
      For twice continuously differentiable functions $\function{f,g}{\setReal^+}{\setComplex}$ we define the Wronskian determinant.
      \[
        W(f,g)\define
        \begin{vmatrix}
          f & g \\
          f' & g'
        \end{vmatrix}
        = fg' - gf'
      \]
      Then the derivative of this functional determinant is given by the following expression.
      \[
        \derivative W(f,g)
        \begin{aligned}[t]
          &= f'g' + fg'' - g'f' - gf'' \\
          &= fg'' - gf''
        \end{aligned}
      \]
      Inserting now $g_l$ and $h_l$ we can prove the proposition.
      \[
        \begin{aligned}[t]
          \derivative &W(g_l,h_l)(r) \\
          &= g_l(r) h_l''(r) - h_l(r) g_l''(r) \\
          &=
            \begin{aligned}[t]
              &g_l(r) h_l(r) \boxBrackets{U_2(r) + \frac{l(l+1)}{r^2} - k^2} \\
              &-h_l(r) g_l(r) \boxBrackets{U_1(r) + \frac{l(l+1)}{r^2} - k^2} \\
            \end{aligned}
          \\
          &= g_l(r) h_l(r) \boxBrackets{U_2(r) - U_1(r)}
        \end{aligned}
      \]
      \hfill$\square$

      \paragraph{Phase Shift Difference:}
      The integral for the phase shift difference can be computed straightforward by using the results above and the fundamental theorem of calculus.
      \[
        \begin{aligned}
          -\frac{2m}{\hbar^2 k} &\integral{0}{\infty}{g_l(r) h_l(r) \boxBrackets{V_2(r) - V_1(r)}}{r} \\
          &= -\frac{1}{k} \integral{0}{\infty}{g_l(r) h_l(r) \boxBrackets{U_2(r) - U_1(r)}}{r} \\
          &= -\frac{1}{k} \integral{0}{\infty}{\derivative W(g_l,h_l)(r)}{r} \\
          &= -\frac{1}{k} \appendValue{W(g_l,h_l)(r)}{r=0}^\infty \\
          &= -\frac{1}{k} \appendValue{\boxBrackets{g_l(r)h_l'(r) - h_l(r) g_l'(r)}}{r=0}^\infty \\
        \end{aligned}
      \]
      Because the radial functions have to be regular at the origin, we are assuming the following conditions.
      \[
        g_l(0) = h_l(0) = 0
      \]
      We already know about the asymptotic behavior of $g_l$ and $h_l$.
      Choose $r\in\setReal^+$ to be near infinity.
      \[
        g_l(r) \converges \sin\roundBrackets{ kr -\frac{π}{2}l + δ_l^{(1)} }
      \]
      \[
        h_l(r) \converges \sin\roundBrackets{ kr -\frac{π}{2}l + δ_l^{(2)} }
      \]
      Therefore we can derive the asymptotic behavior of $h_l'$ and $g_l'$ as well by computing the derivative of the asymptotic function.
      \[
        g_l'(r) \converges k \cos\roundBrackets{ kr -\frac{π}{2}l + δ_l^{(1)} }
      \]
      \[
        h_l'(r) \converges k \cos\roundBrackets{ kr -\frac{π}{2}l + δ_l^{(2)} }
      \]
      We will now insert these ideas into the calculation of the integral.
      Additionally, we use an addition theorem to finish the proof.
      \[
        \begin{aligned}
          -\frac{2m}{\hbar^2 k} &\integral{0}{\infty}{g_l(r) h_l(r) \boxBrackets{V_2(r) - V_1(r)}}{r} \\
          &= -\frac{1}{k} \lim_{r\to\infty}\boxBrackets{g_l(r)h_l'(r) - h_l(r) g_l'(r)} \\
          &= -\frac{1}{k} \lim_{r\to\infty}
            \begin{aligned}[t]
              &\left[ \sin\roundBrackets{ kr -\frac{π}{2}l + δ_l^{(1)} } \right.\\
              &\quad \cdot k \cos\roundBrackets{ kr -\frac{π}{2}l + δ_l^{(2)} } \\
              &- \sin\roundBrackets{ kr -\frac{π}{2}l + δ_l^{(2)} } \\
              &\quad \left. \cdot k \cos\roundBrackets{ kr -\frac{π}{2}l + δ_l^{(1)} } \right]
            \end{aligned}
          \\
          &= \lim_{r\to\infty}
            \begin{aligned}[t]
              &\left[ \sin\roundBrackets{ kr -\frac{π}{2}l + δ_l^{(2)} } \right.\\
              &\quad\cdot \cos\roundBrackets{ kr -\frac{π}{2}l + δ_l^{(1)} } \\
              &- \sin\roundBrackets{ kr -\frac{π}{2}l + δ_l^{(1)} } \\
              &\quad \left. \cdot \cos\roundBrackets{ kr -\frac{π}{2}l + δ_l^{(2)} } \right]
            \end{aligned}
          \\
          &= \lim_{r\to\infty} \sin
              \begin{aligned}[t]
                &\left[ \roundBrackets{kr -\frac{π}{2}l + δ_l^{(2)}} \right. \\
                &\left. - \roundBrackets{kr -\frac{π}{2}l + δ_l^{(1)}} \right]
              \end{aligned}
          \\
          &= \lim_{r\to\infty} \sin \roundBrackets{δ_l^{(2)} - δ_l^{(1)}} \\
          &= \sin \roundBrackets{δ_l^{(2)} - δ_l^{(1)}}
        \end{aligned}
      \]
      \hfill$\square$
    \end{multicols}
  % section problem_19 (end)

  \newpage

  \section*{Problem 20: Identical Particles in an Infinite Well} % (fold)
  \label{sec:problem_20}
    \begin{multicols}{2}
      \paragraph{Preliminaries:}
      Let $\mathscr{L}^2(\setReal)$ be the space of square-integrable functions with domain $\setReal$ and Lebesgue-measure λ.
      \[
        \mathscr{L}^2(\setReal) \define \set{\function{f}{\setReal}{\setComplex}}{\integral{\setReal}{}{\absolute{f}^2}{λ} < \infty}
      \]
      We define the typical scalar product for such a space as follows.
      \[
        \function{\bracket{\cdot}{\cdot}}{\mathscr{L}^2\times\mathscr{L}^2}{\setComplex}
        \separate
        \bracket{f}{g} \define \integral{\setReal}{}{\bar{f}g}{λ}
      \]
      The Hilbert space for a single particle with mass $m\in\setReal^+$ in an infinite potential well with size $a\in\setReal^+$ centered at the origin is described by $\roundBrackets{\mathscr{L}^2(\setReal), \bracket{\cdot}{\cdot}}$.
      For convenience, we define the following constants.
      \[
        ω \define \frac{π}{2a}
        \separate
        ε \define \frac{\hbar^2ω^2}{2m}
        \separate
        z \define \frac{1}{\sqrt{a}}
      \]
      With $\ket{n}\in\mathscr{L}^2(\setReal)$, we denote the eigenstates of the Hamiltonian $\function{H}{\mathscr{L}^2(\setReal)}{\mathscr{L}^2(\setReal)}$ with eigenvalues $E_n$ for all $n\in\setNatural$.
      \[
        \ket{n} \define
        \begin{cases}
          z \cos(nω\cdot) &: n=2k-1 \quad \text{for } k\in\setNatural \\
          z \sin(nω\cdot) &: n=2k \quad \text{for } k\in\setNatural
        \end{cases}
      \]
      \[
        E_n \define εn^2
      \]

      \paragraph{Hilbert Space:}
      For three identical particles the following space $\mathscr{H}$ builds a superset of the Hilbert space for these particles.
      \[
        \mathscr{H} \define \mathscr{L}^2(\setReal) \otimes \mathscr{L}^2(\setReal) \otimes \mathscr{L}^2(\setReal) \cong \mathscr{L}^2\roundBrackets{\setReal^3}
      \]
      To define the scalar product for this space we first define the product state for all $\ket{φ_1},\ket{φ_2},\ket{φ_3}\in\mathscr{L}^2(\setReal)$.
      \[
        \ket{φ_1φ_2φ_3} \define \ket{φ_1}\otimes\ket{φ_2}\otimes\ket{φ_3}
      \]
      Please note, that for convenience we will use function overloading to define the scalar product.
      \[
        \function{\bracket{\cdot}{\cdot}}{\mathscr{H}\times\mathscr{H}}{\setComplex}
      \]
      \[
        \bracket{φ_1φ_2φ_3}{ϑ_1ϑ_2ϑ_3} \define \bracket{φ_1}{ϑ_1} \bracket{φ_2}{ϑ_2} \bracket{φ_3}{ϑ_3}
      \]
      This scalar product will also be used for the Hilbert space of the three identical particles, which is possible because its a subset of $\mathscr{H}$.

      All particles are Bosons because of their spin which is equal to zero.
      Therefore $\mathscr{H}$ has to be symmetrized such that all possible wave functions are symmetric.
      We call $\roundBrackets{\mathscr{H}_\mathrm{S},\bracket{\cdot}{\cdot}}$ the Hilbert space of these three identical Bosons.
      \[
        \mathscr{H}_\mathrm{S} \define \set{\ket{φ}\in\mathscr{H}}{\forall π\in\mathrm{S}_3:\ P(π)\ket{φ}=\ket{φ}} \subset \mathscr{H}
      \]
      Here, $\mathrm{S}_3$ is the three-dimensional permutation group and $P(π)$ the related permutation operator for $π\in\mathrm{S}_3$.

      \paragraph{Eigenbasis:}
      The three particles do not interact.
      Hence, the Hamiltonian $\function{H_\mathrm{S}}{\mathscr{H}_\mathrm{S}}{\mathscr{H}_\mathrm{S}}$ for $\mathscr{H}_\mathrm{S}$ can be stated as follows.
      \[
        H_\mathrm{S} \define H\otimes\mathds{1}\otimes\mathds{1} + \mathds{1}\otimes H\otimes\mathds{1} + \mathds{1}\otimes\mathds{1}\otimes H
      \]
      We can now conclude that for all $n_1,n_2,n_3\in\setNatural$ the states $\ket{n_1n_2n_3}$ are eigenstates of $H_\mathrm{S}\vert_{\mathscr{H}}$ with eigenvalue $E_{n_1n_2n_3}$.
      \begin{align*}
        H_\mathrm{S}\ket{n_1n_2n_3}
        &= H\ket{n_1}\otimes\ket{n_2}\otimes\ket{n_3} \\
        &+ \ket{n_1}\otimes H\ket{n_2}\otimes\ket{n_3} \\
        &+ \ket{n_1}\otimes\ket{n_2}\otimes H\ket{n_3} \\
        &= \roundBrackets{E_{n_1} + E_{n_2} + E_{n_3}} \ket{n_1n_2n_3}
      \end{align*}
      \[
        E_{n_1n_2n_3} \define E_{n_1} + E_{n_2} + E_{n_3}
      \]
      Additionally, for all $m_1,m_2,m_3\in\setNatural$ the following equation shows that the states are orthonormal and consequently are building an orthonormal basis of $\mathscr{H}$.
      \begin{align*}
        &\bracket{m_1m_2m_3}{n_1n_2n_3} \\
        &= \bracket{m_1}{n_1}\bracket{m_2}{n_2}\bracket{m_3}{n_3} \\
        &= δ_{m_1n_1}δ_{m_2n_2}δ_{m_3n_3}
      \end{align*}
      But these states are in general not symmetric and may not lie in $\mathscr{H}_\mathrm{S}$ which would violate the condition that the system consists of three identical Bosons.
      Like $\mathscr{H}$ had to be symmetrized, the states also have to be symmetrized.
      For all $n_1,n_2,n_3\in\setNatural$ we define the symmetric states as follows.
      \[
        \ket{n_1n_2n_3}_\mathrm{S}\define \sum_{σ\in\mathrm{S_3}}\ket{n_{σ(1)}n_{σ(2)}n_{σ(3)}}
      \]
      Due to the linear combination and the permutation we are now making sure that the states are building a symmetric eigenbasis of $\mathscr{H}_\mathrm{S}$ with respect to $H_\mathrm{S}$.
      Of course, these symmetric states are not normalized.
      To normalize the states, we choose $n_1,n_2,n_3\in\setNatural$ to be arbitrary.
      \begin{align*}
        &\prescript{}{\mathrm{S}}{\bracket{n_1n_2n_3}{n_1n_2n_3}_\mathrm{S}} \\
        &= \sum_{π,σ\in\mathrm{S}_3} \bracket{n_{π(1)}n_{π(2)}n_{π(3)}}{n_{σ(1)}n_{σ(2)}n_{σ(3)}} \\
        &= \sum_{π,σ\in\mathrm{S}_3} δ_{n_{π(1)}n_{σ(1)}} δ_{n_{π(2)}n_{σ(2)}} δ_{n_{π(3)}n_{σ(3)}}
      \end{align*}
      We have to sum over $3!\cdot3!=36$ terms and at least six of them will be one because the permutations $σ$ and $π$ can be the same.

      \paragraph{Case $n_1\neq n_2\neq n_3 \neq n_1$:}
      Then it is clear that only these six terms can contribute to the actual sum because for every permutation $π\in\mathrm{S}_3$ we compute the result below.
      \begin{align*}
        &\sum_{σ\in\mathrm{S}_3}\bracket{n_{π(1)}n_{π(2)}n_{π(3)}}{n_{σ(1)}n_{σ(2)}n_{σ(3)}} \\
        &= \sum_{σ\in\mathrm{S}_3} δ_{{π(1)}{σ(1)}} δ_{{π(2)}{σ(2)}} δ_{{π(3)}{σ(3)}}
        = 1
      \end{align*}
      The normalization in this case is straightforward.
      \[
        \ket{n_1n_2n_3}_{\tilde{\mathrm{S}}} \define \frac{1}{\sqrt{3!}}\ket{n_1n_2n_3}_\mathrm{S}
      \]

      \paragraph{Case $n_1\neq n_2 = n_3$:}
      If two values are the same then a permutation interchanging those values will not change the actual state.
      This can be stated as follows for all permutations $π\in\mathrm{S}_3$.
      \begin{align*}
        &\sum_{σ\in\mathrm{S}_3}\bracket{n_{π(1)}n_{π(2)}n_{π(3)}}{n_{σ(1)}n_{σ(2)}n_{σ(3)}} \\
        &= \sum_{σ\in\mathrm{S}_3}
          \begin{aligned}[t]
            & \left( δ_{{π(1)}{σ(1)}} δ_{{π(2)}{σ(2)}} δ_{{π(3)}{σ(3)}} \right. \\
            & \left. + δ_{{π(1)}{σ(1)}} δ_{{π(2)}{σ(3)}} δ_{{π(3)}{σ(2)}} \right)
          \end{aligned}
        \\
        &= 2
      \end{align*}
      In this case the normalization factor changes and the normalized state is defined below.
      \[
        \ket{n_1n_2n_3}_{\tilde{\mathrm{S}}} \define \frac{1}{\sqrt{3!\cdot 2!}}\ket{n_1n_2n_3}_\mathrm{S}
      \]

      \paragraph{Case $n_1=n_2=n_3$:}
      If all of these three values are the same then every term in the sum has to be one because for every permutation $π\in\mathrm{S}_3$ we can conclude the following.
      \begin{align*}
        &\sum_{σ\in\mathrm{S}_3}\bracket{n_{π(1)}n_{π(2)}n_{π(3)}}{n_{σ(1)}n_{σ(2)}n_{σ(3)}} \\
        &= \sum_{σ\in\mathrm{S}_3}\sum_{κ\in\mathrm{S}_3} δ_{{π(1)}{σ(κ(1))}} δ_{{π(2)}{σ(κ(2))}} δ_{{π(3)}{σ(κ(3))}} \\
        &= 6
      \end{align*}
      Of course, this is a direct consequence of the following equation for all $σ\in\mathrm{S}_3$.
      \[
        \ket{n_1n_2n_3} = \ket{n_{σ(1)}n_{σ(2)}n_{σ(3)}}
      \]
      Hence, the normalization in this case can formulated as follows.
      \[
        \ket{n_1n_2n_3}_{\tilde{\mathrm{S}}} \define \frac{1}{\sqrt{3!\cdot 3!}}\ket{n_1n_2n_3}_\mathrm{S}
      \]

      \paragraph{General Normalization:}
      The results obtained here can be generalized.
      Define $p_n\in\setNatural_0$ to be the count of particles in the state $\ket{n}$ for all $n\in\setNatural$.
      Because we have three particles the following equation is fulfilled.
      \[
        \sum_{n\in\setNatural} p_n = 3
      \]
      Particles in the same state can be interchanged.
      There are $p_n!$ possible permutations for every $n\in\setNatural$ such that the state is not changed.
      Thus, for every permutation $π\in\mathrm{S}_3$ the following holds.
      \[
        \sum_{σ\in\mathrm{S}_3}\bracket{n_{π(1)}n_{π(2)}n_{π(3)}}{n_{σ(1)}n_{σ(2)}n_{σ(3)}}
        = \prod_{n\in\setNatural} p_n!
      \]
      With this we could have also defined the following general normalization rule.
      \[
        \ket{n_1n_2n_3}_{\tilde{\mathrm{S}}} \define \frac{1}{\sqrt{3!\prod_{n\in\setNatural}p_n!}} \ket{n_1n_2n_3}_\mathrm{S}
      \]
      The set $\set{\ket{n_1n_2n_3}_{\tilde{\mathrm{S}}}}{n_1,n_2,n_3\in\setNatural}$ is therefore building an orthonormal eigenbasis of $\mathscr{H}_\mathrm{S}$ with respect to $H_\mathrm{S}$.
      Please note, that the states are symmetric and therefore the order of $n_1$, $n_2$ and $n_3$ can be ignored.
      It suffices to describe the symmetric state by the numbers $(p_n)_{n\in\setNatural}$.

      \paragraph{Lowest Energy States:}
      As shown above for a normalized eigenstate $\ket{n_1n_2n_3}_{\tilde{\mathrm{S}}}$ of the Hamiltonian $H_\mathrm{S}$ with have the energy $E_{n_1n_2n_3}$ as an eigenvalue for all $n_1,n_2,n_3\in\setNatural$.
      Therefore we can find the five states with the lowest energy by direct computation.
      \begin{align*}
        &\ket{111}_{\tilde{\mathrm{S}}}: & E_{111} &= 3ε \\
        &\ket{112}_{\tilde{\mathrm{S}}}: & E_{112} &= 6ε \\
        &\ket{122}_{\tilde{\mathrm{S}}}: & E_{122} &= 9ε \\
        &\ket{113}_{\tilde{\mathrm{S}}}: & E_{113} &= 11ε \\
        &\ket{222}_{\tilde{\mathrm{S}}}: & E_{222} &= 12ε
      \end{align*}
    \end{multicols}
  % section problem_20 (end)
\end{document}